Location and Scale-Invariant Power Transformations for Transforming Data to

### 2025-02-26 manuscript_rmarkdown/manuscript.Rmd
WIP on additional ML models.

We used 285 datasets from the Penn Machine Learning Benchmarks collection [@Romano2022-gq].
In this collection, 122 datasets correspond to regression tasks and 163 datasets
to classification tasks. Using the familiar auto-machine learning library [@Zwanenburg2021-so] (version 1.5.0),
each dataset was used to train a model for each of 16 process configurations.
Each process configuration specifies the learner (generalised linear model or
random forest), transformation method (none, conventional Yeo-Johnson, robust
invariant Yeo-Johnson, robust invariant Yeo-Johnson with empirical central
normality test (rejecting transformations with $p \leq 0.01$), and normalisation method (none,
$z$-standardisation), yielding 16 distinct configurations.
Before each experiment, each dataset was randomly split into a training (70%)
and holdout test (30%) set five times. Thus, a total of 22800 models were
created. Each model was then evaluated using the holdout test set using one of
two metrics, i.e. the root relative squared error (RRSE) for regression tasks and the
area under the receiver operating characteristic curve (AUC) for classification tasks.

@@

* very easy: $\text{AUC} \geq 0.90$ or $\text{RRSE} \leq 0.10$ (87 datasets)
* easy: $0.90 > \text{AUC} \geq 0.80$ or $0.30 \geq \text{RRSE} > 0.10$ (46 datasets)
* intermediate: $0.80 > \text{AUC} \geq 0.70$ or $0.60 \geq \text{RRSE} > 0.30$ (72 datasets)
* difficult: $0.70 > \text{AUC} \geq 0.60$ or $0.80 \geq \text{RRSE} > 0.60$ (60 datasets)
* very difficult: $0.60 > \text{AUC} \geq 0.50$ or $1.00 \geq \text{RRSE} > 0.80$ (12 datasets)
* unsolvable: $\text{AUC} < 0.50$ or $\text{RRSE} > 1.00$ (8 datasets)

@@

# Convert to data matrix. First aggregate within each experiment because the
# Wilcoxon signed rank test is not for repeated measurements.
data <- data.table::dcast(
  data,
  dataset ~ learner, value.var="value_rank",
  fun.aggregate = mean
)

h_learner <- stats::wilcox.test(
    x = data[["random_forest_ranger"]],
    y = data[["glm"]],
    paired = TRUE,
    alternative = "two.sided"
)
```

@@

Here we focus on the subset of 232 datasets that contain numeric features. 
Considering single process parameters, the choice of learner (Wilcoxon signed rank test: $p < 10^{-8})$) and normalisation method (Wilcoxon signed rank test: $p = 0.007$) had a significant impact (at significance level $p = 0.050$), but transformation method (Friedman test: $p = 0.054$) did not.

>>>>

We used 231 datasets containing at least one numeric feature from the Penn Machine Learning Benchmarks collection [@Romano2022-gq].
In this collection, 114 datasets correspond to regression tasks and 117 datasets
to classification tasks. Using the familiar auto-machine learning library [@Zwanenburg2021-so] (version 1.5.0),
each dataset was used to train a model for each of 32 process configurations.
Each process configuration specifies the learner (generalised linear model, L1-regularised linear models (Lasso), gradient boosted linear model, or
random forest), transformation method (none, conventional Yeo-Johnson, robust
invariant Yeo-Johnson, robust invariant Yeo-Johnson with empirical central
normality test (rejecting transformations with $p \leq 0.01$), and normalisation method (none,
$z$-standardisation), yielding 32 distinct configurations.
Before each experiment, each dataset was randomly split into a training (70%)
and holdout test (30%) set five times. Thus, a total of 27720 models were
created. Each model was then evaluated using the holdout test set using one of
two metrics, i.e. the root relative squared error (RRSE) for regression tasks and the
area under the receiver operating characteristic curve (AUC) for classification tasks.

@@

* very easy: $\text{AUC} \geq 0.90$ or $\text{RRSE} \leq 0.10$ (59 datasets)
* easy: $0.90 > \text{AUC} \geq 0.80$ or $0.30 \geq \text{RRSE} > 0.10$ (40 datasets)
* intermediate: $0.80 > \text{AUC} \geq 0.70$ or $0.60 \geq \text{RRSE} > 0.30$ (50 datasets)
* difficult: $0.70 > \text{AUC} \geq 0.60$ or $0.80 \geq \text{RRSE} > 0.60$ (32 datasets)
* very difficult: $0.60 > \text{AUC} \geq 0.50$ or $1.00 \geq \text{RRSE} > 0.80$ (49 datasets)
* unsolvable: $\text{AUC} < 0.50$ or $\text{RRSE} > 1.00$ (1 dataset)

@@

# Convert to data matrix. First aggregate within each experiment because the
# Friedman test is not for repeated measurements.
data <- data.table::dcast(
  data,
  dataset ~ learner, value.var="value_rank",
  fun.aggregate = mean
)

h_learner <- friedman.test(as.matrix(data[, "dataset" := NULL]))

# h_learner <- stats::wilcox.test(
#     x = data[["random_forest_ranger"]],
#     y = data[["glm"]],
#     paired = TRUE,
#     alternative = "two.sided"
# )
```

@@

Considering single process parameters, the choice of learner (Friedman test: $p < 10^{-8})$), normalisation method (Wilcoxon signed rank test: $p < 10^{-5}$), and transformation method $p = 0.006$, all had a significant impact (at significance level $p = 0.05$).
